---
title: "R Notebook"
output: html_notebook
---

# The apply family
lapply

Aplly a function to each element of a list and return a list

```{r}
lapply(1:10, function(x) x^2)

lapply(1:10, function(x) x-2)
lapply(1:10, function(x) {x-23})

```

sapply return a vector 

```{r}
sapply (1:10, function(x) x^2)

```

Group and summarizing

mean() and sum() are only tow of the functions R provides for summarizing a collection of numbers.
```{r}
library(gapminder)
library(dplyr)
gapminder %>%
  summarise(mean_lifeExp = median(lifeExp))
```

Instead of grouping just by year, or just by continent, youll now froup by both continent and year to summarize within each.

Find the median life expectancy and maximun GDP per capita within each combination and year, saving them into medianLifeExp and maxGdpPercap respectively..

```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarise(medianLifeExp = median(lifeExp),
            maxGdpPercap = max(gdpPercap))
```

```{r}
by_continent <- gapminder %>%
  filter(year == 1952) %>%
  group_by(continent) %>%
  summarise(medianGdpPercap = median(gdpPercap))
```

```{r}
library(ggplot2)
ggplot(by_continent, aes(x = continent, y = medianGdpPercap)) +
  geom_bar(stat = "identity")
```

```{r}
library(tidyverse)
counties <- read.csv("counties.csv")
glimpse(counties)

select(counties, state, county, population, men, women)
```


